---
title: "AIMS weather time series data sets"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{AIMS weather time series data sets}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This example assumes you have obtained an [AIMS Data Platform API Key](https://open-aims.github.io/data-platform/key-request). Further details about API Key implementation and usage can be found [here](../index.html).

To interact locally with this example in [RStudio](https://www.rstudio.com/), click on the source code at the top of this page, then click on `Raw`, and simply copy and paste the code to a new [.Rmd](https://rmarkdown.rstudio.com/) file created in [RStudio](https://www.rstudio.com/).

Let's start by loading some packages that we are going to need down the track:

```{r, eval = FALSE}
library(plyr)
library(dplyr)
library(ggplot2)
```

Once you have obtained your API Key, please consider following the recommended guidelines as explained [here](../index.html). If you successfully placed your API Key permanently to the `.Renviron`, then just set the object `my_api_key` to `NULL` in the chunk below. If for some reason you are having difficulty placing your API Key permanently to the `.Renviron`, then just paste it to the object `my_api_key`.

```{r, echo = FALSE, message = FALSE}
library(plyr)
library(dplyr)
library(ggplot2)
library(dataaimsr)
my_api_key <- "x78NRh0XrH9Irtju5OoOdTYBhG0pklw2w6o7eos6"
```

```{r, eval = FALSE}
# set my_api_key to NULL if successfully placed in .Renviron
# paste your key where it says api-key-for-r-notebook
my_api_key <- "api-key-for-r-notebook"
```

## Install R Client

`dataaimsr` can be installed and loaded directly from [GitHub](https://github.com/open-aims/dataaimsr) using the following code:

```{r, eval = FALSE}
install.packages("devtools")
devtools::install_github("https://github.com/open-AIMS/dataaimsr")
library(dataaimsr)
```

## Discover a dataset

Suppose we want to query a data series and plot it then the procedure might be:

1. Examine documentation and establish query filters
2. Perform data download using `aims_data`
3. Use the Rpackage `ggplot2` to create a time-series chart

There are two types of data programs currently available through the [AIMS Data Platform API](https://open-aims.github.io/data-platform): [Weather](https://open-aims.github.io/data-platform/weather), which contains multiple datasets for different variables (e.g. Air Temperature, Air Pressure, Chlorophyll, and many others), and [Sea Water Temperature Loggers](https://open-aims.github.io/data-platform/temperature-loggers), which contains multiple datasets of (you guessed it!) sea water temperature at different sites and water depths. They are queried via unique DOI identifiers.

For the purposes of this example, we will query the [AIMS Weather](https://open-aims.github.io/data-platform/weather). To explore this data program, we use the function `filter_values`. This function takes a DOI and a given parameter filter, and returns all the available information regarding the filter. For the Weather data sets, a list of available filter parameters can be found [here](https://open-aims.github.io/data-platform/weather/index). We recommend exploring the parameter `series`:

```{r}
weather_doi  <-  aims_data_doi("weather")
filter_values(weather_doi, filter_name = "series") %>%
  head()
```

The benefits to choosing a data `series` is that it comes from one location and variable type, making the data easy to plot. If we did not choose a data series from the [AIMS Weather](https://open-aims.github.io/data-platform/weather) program, we would have to specify additional parameters to ensure the data is as expected.

Finally our values and filters might look like the following:

Variable  | Value                  | Description
----------|------------------------|-------------------------------------------------------
doi       | 10.25845/5c09bf93f315d | DOI for AIMS Weather found [here](https://open-aims.github.io/data-platform)
series    | 64                     | Found [here](https://open-aims.github.io/data-platform/weather/series), Davies Reef Air Temperature data series
from-date | '2018-01-01'           | We want to start charting on 1/1/2018
thru-date | '2018-01-10'           | We are plotting 10 days of data

## Query and Plot Dataset

After deciding on query parameters, we plug the series id into a `aims_data` function:

```{r, message = FALSE}
davies <- aims_data(weather_doi,
                    api_key = my_api_key,
                    filters = list("series" = 64,
                                   "from-date" = "2018-01-01",
                                   "thru-date" = "2018-01-10"))
```

The object `davies` is a list containing four elements:  
- `metadata` a doi link containing the metadata record for the data series  
- `citation` the citation information for the particular dataset 
- `links` the link returning the data query  
- `data` an output `data.frame`  

We can even visually compare multiple series at once. For instance, let's compare the air temperature data from Davies Reef and Bramble Cay for the same period of time:

```{r, message = FALSE}
target_series <- c("Davies Reef" = 64, "Bramble Cay" = 87929)
results <- plyr::llply(target_series, function(series_number, my_api_key, ...) {
  aims_data(weather_doi, api_key = my_api_key, 
            filters = list("series" = series_number, ...))
}, my_api_key = my_api_key, "from-date" = "2018-01-01", "thru-date" = "2018-01-10")

weather_data <- plyr::ldply(results, "[[", "data")

ggplot(weather_data) +
  geom_line(aes(x = time, y = raw_value, colour = site_name)) +
  labs(x = "Date",
       y = "Air temperature (˚C)",
       colour = "Site",
       title = "AIMS",
       subtitle = "Weather stations") +
  theme_bw() +
  theme(axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        legend.position = "bottom")
```

## Bibliography

```{r, message = FALSE}
plyr::ldply(results, "[[", "citation") %>%
  dplyr::rename("citation" = "[[") %>%
  dplyr::select(citation) %>%
  unlist %>%
  unname
```
